---
title: "FRS_project"
author: "Elena"
date: '2022-08-19'
output: html_document
---
Data to compile from FRS:

 * Facility Name  	
 * Street Address	
 * City  
 * Zip  	
 * County	
 * State 	
 * Latitude	
 * Longitude 	
 * FRS ID	X 
 * NAICS Code	
 * SIC Code 
 

To upload the data from FRS go to https://www.epa.gov/frs/epa-state-combined-csv-download-files click on California on the map.
Zip file (200 Mb) will be downloaded automatically. Or [click here](https://ordsext.epa.gov/FLA/www3/state_files/state_combined_ca.zip).

The purpose of this code is to: 

**(1)** extract data according to above list by: 

 **(1.1)** performing filtering according to selected NAICS (31-33)/SIC(20-39) codes, 
 
 **(1.2)** joining the tables by Registry Id (or FRS Id - used interchangeably),
 
 **(1.3)** cleaning by removing (i.e., saving separately) missing data for further investigation.

The result is three tables: 
 - *FRS_bothCodes* (open to suggestion for better naming!): 46763 lines, 36238 Unique Facilities 

 - *FRS_noCoords*: 7581 lines, 7279 Unique Facilities 

 - *FRS_corrupted_county*: 1339 lines, 1321 Unique Facilities.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) #for tidy data
library(readr) # for efficient read of files
library(naniar) # for dealing with NAs
library(janitor) #for automatically cleaning names from upeer case to lower
library(stringr) #for cleaning the strings

```

#Upload file with FRS Facilities and their Addresses
```{r CA_FACILITY_FILE.CSV}
FRS_Address <- read.csv("CA_FACILITY_FILE.CSV") 
FRS_Address <- clean_names(FRS_Address)
names(FRS_Address)

#remove unnecessary columns. leave frsId, company name, address, city, county, state,zip, lat, long
FRS_Address <- FRS_Address[, c(2:4, 6,7, 9, 12, 27, 28)] # total number of addressees is 654419; all lines in the df are unique

#length(unique(FRS_Address$registry_id))
#sum(duplicated(FRS_Address)) #0
#vis_miss(FRS_Address, warn_large_data = FALSE) #use only if you have memory (takes up to 3 Gb with this data set)

```

In address file we have 32.23% of missing coordinates (for ALL facilities), meaning around 1/3 of the FRS data do not have coordinates responded to the Facility Address. Later, we will see that percentage of missing coordinates for the subset of interest is much lower.
I will subset FRS_noCoords with codes of interest later here. This way we can perform geocoding.
Thus, there must be two final files: 1) full file with all facilities of interest, 2) with facilities where coordinates are absent.


#Prepare Facilities with NAICS of interest
```{r CA_NAICS_FILE.csv}
#read NAICS codes
FRS_NAICS <- read.csv("CA_NAICS_FILE.csv") #number of rows = 468732
FRS_NAICS <- janitor::clean_names(FRS_NAICS) #lower case all column headers
names(FRS_NAICS)

#leave columns: frsId, naics code, code description
FRS_NAICS <- FRS_NAICS[, c(1,5,7)]

#check how many NAs
#vis_miss(FRS_NAICS, warn_large_data = FALSE) #no NAs

#length(unique(FRS_NAICS$registry_id)) #unique FRS_id = 370020 => meaning that some FRS_id report to programs by several NAICS 
#sum(duplicated(FRS_NAICS)) #56063 duplicates detected (for FRS_id, naics code, description) after I removed other columns

#drop duplictes
FRS_NAICS <- distinct(FRS_NAICS) #412669 left after removing duplicates
412669+56063 #=468732

#FRSwithNAICS_Frequency <- as.data.frame(table(FRS_NAICS$registry_id)) #some FRS Id report with up to 10 different NAICS codes
412669+56063 #=468732

#leave only manufacturing facilities (31-33)
FRS_NAICS <- FRS_NAICS %>%  filter(grepl("^31",naics_code) | grepl("^32", naics_code)| grepl("^33", naics_code)) #after filter, only 46617 lines left: 38001 unique facilities left

#remove naics codes that were rejected during the meeting
FRS_NAICS <- FRS_NAICS %>%  filter(!naics_code %in% c(311, 
3111, 31111, 311111, 311119, 3112, 31121, 311211, 311212, 311213, 31122, 311221, 311224, 311225, 31123, 311230, 3113, 31131, 311313, 311314, 31134, 311340, 31135, 311351, 311352, 3114, 31141, 311411, 311412, 31142, 311421, 311422, 311423, 3115, 31151, 311511, 
311512, 311513, 311514, 31152, 311520, 3116, 31161, 311611, 311612, 311613, 311615, 3117, 31171, 311710, 3118, 31181, 311811, 311812, 
311813, 31182, 311821, 311824, 31183, 311830, 3119, 31191, 311911, 311919, 31192, 311920, 31193, 311930, 31194, 311941, 311942, 31199, 311991, 311999, 312, 3121, 31211, 312111, 312112, 312113, 31212, 312120, 31213, 312130, 31214, 312140, 32312, 323120, 3253, 32532, 
325320, 325412, 325413, 325414, 3391, 33911, 339112, 339113, 339114, 339116)) 

#length(unique(FRS_NAICS$registry_id)) #32379 unique facilities left
```

## Facilities with apropriate NAICS and Address
```{r Join Address with NAICS of interest}
FRS_address_naics <- inner_join(FRS_Address, FRS_NAICS, by = "registry_id") #39687 lines

#sum(duplicated(FRS_Address)) #0
#vis_miss(FRS_address_naics, warn_large_data = FALSE)
#length(unique(FRS_address_naics$registry_id)) #32379 unique facilities

```

Out of 654419 frs_id, 370020 facilities have NAICS (this statement migrated from previous version of transforming FRS data).
Out of 370020 facilities with NAICS, 32379 facility of interest.

#Prepare Facilities with SIC of interest
```{r CA_SIC_FILE.CSV}
FRS_SIC <- read.csv("CA_SIC_FILE.CSV") #total of 168253 lines

FRS_SIC <- clean_names(FRS_SIC)
names(FRS_SIC)

FRS_SIC <- FRS_SIC[, c(1,5,7)]

#length(unique(FRS_SIC$registry_id)) #118069 has unique frs id
#sum(duplicated(FRS_SIC))  #out of 168253, 44330 are duplicates

FRS_SIC <- distinct(FRS_SIC) #123923 are total line left after removing duplicates
#vis_miss(FRS_SIC, warn_large_data = FALSE)

#filter for the office of manufacturing SIC (20-39)
FRS_SIC <- FRS_SIC %>%  filter(grepl("^2",sic_code) | grepl("^3",sic_code)) #30517 sic codes from office of manufacturing is left
#???

FRS_SIC <- FRS_SIC %>% filter(!sic_code %in% c(2011, 2013, 2015, 2021, 2022, 2023, 2024, 2026, 2032, 2033, 2034, 2035, 2037, 2038, 2041, 2043, 2044, 2045, 2046, 2047, 2048, 2051, 2052, 2053, 2061, 2062, 2063, 2064, 2066, 2067, 2068, 2074, 2075, 2076, 2077, 2079, 
2082, 2083, 2084, 2085, 2086, 2087, 2091, 2092, 2095, 2096, 2097, 2098, 2099, 2499, 2599, 2789, 2791, 2796, 2834, 2835, 2836, 2879, 2899, 3069, 3089, 3821, 3829, 3841, 3842, 3843, 3851, 2411, 2711, 2721, 2731, 2741, 2771, 3652, 3721, 3732))
#23332 sic lines; 21607 unique facilities

```

## Facilities with apropriate SIC and Address
```{r Join Address with SIC (20-39)}
FRS_address_sic <- inner_join(FRS_Address, FRS_SIC, by = "registry_id")
#sum(duplicated(FRS_address_sic)) #0

#vis_miss(FRS_address_sic, warn_large_data = FALSE)
```

#Merging NAICS and SIC
```{r Join NAICS and SIC; preserve data with no Coords}
length(unique(FRS_address_naics$registry_id)) #32379 unique frs
length(unique(FRS_address_sic$registry_id))#21607 unique frs

FRS_bothCodes <- full_join(FRS_address_naics, FRS_address_sic, 
                           by = c("registry_id", "primary_name", 
                                  "location_address", "city_name", 
                                  "county_name", "postal_code",
                                  "latitude83", "longitude83", "state_code")) #54345 lines; #this is a Master list by Elena

names(FRS_bothCodes) <- c("frs_id", "primary_name", "facility_address", "city", "county", "state", "ZIP", "lat","long","NAICS","NAICS_description", "SIC", "SIC_description") 

#vis_miss(FRS_bothCodes, warn_large_data = FALSE)

#subset data with absent coordinate for later analysis
FRS_noCoords <- subset(FRS_bothCodes, 
                       (is.na(FRS_bothCodes[,8])) & (is.na(FRS_bothCodes[,9]))) 
#(col 8 and 9 are lat and long accordingly)
length(unique(FRS_noCoords$frs_id)) #out of facilities with codes of interest (NAICS or SIC), 7279 facilities does not have coordinates

#FRS_bothCodes <- FRS_bothCodes %>%  drop_na(lat) #46764 lines with coords left
length(unique(FRS_bothCodes$frs_id)) #43518 unique facilities
```

```{r Cleaning strings in Master file}

#remove NA's from NAICS/SIC and their description
FRS_bothCodes$NAICS[is.na(FRS_bothCodes$NAICS)] <- 0
FRS_bothCodes$SIC[is.na(FRS_bothCodes$SIC)] <- 0
FRS_bothCodes$NAICS_description[is.na(FRS_bothCodes$NAICS_description)] <- "missing"
FRS_bothCodes$SIC_description[is.na(FRS_bothCodes$SIC_description)] <- "missing"

#FRS_bothCodes$facility_address[is.na(FRS_bothCodes$facility_address)] <- "missing"
noaddressE <- filter(FRS_bothCodes, facility_address == "")


FRS_bothCodes$county <- str_to_title(FRS_bothCodes$county)
FRS_bothCodes$city <- str_to_title(FRS_bothCodes$city)
FRS_bothCodes$primary_name <- str_to_title(FRS_bothCodes$primary_name)
FRS_bothCodes$facility_address <- str_to_title(FRS_bothCodes$facility_address)
FRS_bothCodes$NAICS_description <- str_to_title(FRS_bothCodes$NAICS_description)
FRS_bothCodes$SIC_description <- str_to_title(FRS_bothCodes$SIC_description)
```


```{r Cleaning counties and zip-codes}

unique(FRS_bothCodes$county)
FRS_bothCodes$county <- str_replace_all(FRS_bothCodes$county, pattern = "\\sCou\\S+", replacement ="")

county_infrs<- as.data.frame(table(FRS_bothCodes$county))

county_list <- c('Alameda', 'Alpine', 'Amador', 
            'Butte', 
            'Calaveras', 'Colusa', 'Contra Costa', 
            'Del Norte', 
            'El Dorado', 
            'Fresno', 
            'Glenn', 
            'Humboldt', 
            'Imperial', 'Inyo', 
            'Kern', 'Kings', 
            'Lake', 'Lassen', 'Los Angeles', 
            'Madera', 'Marin', 'Mariposa', 'Mendocino', 'Merced', 'Modoc', 'Mono', 'Monterey', 
            'Napa', 'Nevada', 
            'Orange', 
            'Placer', 'Plumas', 
            'Riverside', 
            'Sacramento', 'San Benito', 'San Bernardino', 'San Diego', 'San Francisco', 'San Joaquin', 'San Luis Obispo', 'San Mateo', 'Santa Barbara', 'Santa Clara', 'Santa Cruz', 'Shasta', 'Sierra', 'Siskiyou', 'Solano', 'Sonoma', 'Stanislaus', 'Sutter', 
            'Tehama', 'Tuolumne', 'Trinity', 'Tulare', 
            'Ventura', 
            'Yolo', 'Yuba')
check_county <-county_infrs %>%  mutate(check = ifelse(county_infrs$Var1 %in% county_list, "Valid", "Not valid"))
#the table above will help to see what county names are misspelled and how many of those are present in table

#La is used only once; after verifiying the address with google, La is fixed
FRS_bothCodes$county[FRS_bothCodes$county=="La"] <- "Los Angeles"

#missing counties and Union must be investigated:

#get rid of blanks substituting them by "missing"
FRS_bothCodes$county[FRS_bothCodes$county==""] <- "missing"
#create a df with corrupted counties - *must be investigated*. #missing counties are not removed from the Master table. 
FRS_corrupted_county <- FRS_bothCodes %>% 
  filter(county=="missing" | county == "Union")
#remove "Union" from county. Apparently the facility is not in state. Must be investigated.
FRS_bothCodes <- filter(FRS_bothCodes, county != "Union") 

FRS_bothCodes$ZIP <- substr(FRS_bothCodes$ZIP,1,5)
```

```{r Save master files}
write.csv(FRS_bothCodes, "FRS_facility_codes.csv")
write.csv(FRS_noCoords, "EG_FRSbothCodes_OnlyMissingCoords.csv")

```
both files are combined into excel workbook in teams https://cadtsc.sharepoint.com/sites/spwp-ChemPET/_layouts/15/Doc.aspx?OR=teams&action=edit&sourcedoc={44AB631A-6C40-40BE-88D8-05A13406D6C5}

#Messy code but it will make sense!
```{r from Jess raw data - may ignore}
list.files()

#raw file that Jess queried. What queries did she apply?
Jess<- read.csv("FRS_31_33_raw.csv") #71827 lines

names(Jess)
Jess <- Jess[, c(1:3, 6:11, 17,18)]
Jess <- distinct(Jess) #50781 

#subset only data that have NAICS codes
Jess_naics <- Jess %>%  subset(NAICS_CODE != 0) #38934

#filter for naics of interest on the code
Jess_naics <- Jess_naics %>%  filter(!NAICS_CODE %in% c(311, 
3111, 31111, 311111, 311119, 3112, 31121, 311211, 311212, 311213, 31122, 311221, 311224, 311225, 31123, 311230, 3113, 31131, 311313, 311314, 31134, 311340, 31135, 311351, 311352, 3114, 31141, 311411, 311412, 31142, 311421, 311422, 311423, 3115, 31151, 311511, 
311512, 311513, 311514, 31152, 311520, 3116, 31161, 311611, 311612, 311613, 311615, 3117, 31171, 311710, 3118, 31181, 311811, 311812, 
311813, 31182, 311821, 311824, 31183, 311830, 3119, 31191, 311911, 311919, 31192, 311920, 31193, 311930, 31194, 311941, 311942, 31199, 311991, 311999, 312, 3121, 31211, 312111, 312112, 312113, 31212, 312120, 31213, 312130, 31214, 312140, 32312, 323120, 3253, 32532, 
325320, 325412, 325413, 325414, 3391, 33911, 339112, 339113, 339114, 339116)) #32917 unique

#subset only data that do not have NAICS (meaning data only have SIC codes)
Jess_sic <- Jess %>%  subset(NAICS_CODE == 0) #11847

Jess_sic <- Jess_sic %>% filter(!SIC_CODES %in% c(2011, 2013, 2015, 2021, 2022, 2023, 2024, 2026, 2032, 2033, 2034, 2035, 2037, 2038, 2041, 2043, 2044, 2045, 2046, 2047, 2048, 2051, 2052, 2053, 2061, 2062, 2063, 2064, 2066, 2067, 2068, 2074, 2075, 2076, 2077, 2079, 
2082, 2083, 2084, 2085, 2086, 2087, 2091, 2092, 2095, 2096, 2097, 2098, 2099, 2499, 2599, 2789, 2791, 2796, 2834, 2835, 2836, 2879, 2899, 3069, 3089, 3821, 3829, 3841, 3842, 3843, 3851, 2411, 2711, 2721, 2731, 2741, 2771, 3652, 3721, 3732)) #7843 unique

jess_bothcosedes <- rbind(Jess_naics, Jess_sic) #39981 unique

elena_bothcodes <- FRS_bothCodes

names(elena_bothcodes)
elena_bothcodes <- elena_bothcodes[,c (1:9)]
elena_bothcodes <- distinct(elena_bothcodes)

length(unique(jess_bothcosedes$REGISTRY_ID)) #39981
length(unique(elena_bothcodes$frs_id)) #43517

check_by_join <- full_join(elena_bothcodes, jess_bothcosedes, by = c("frs_id" = "REGISTRY_ID"))

elena_not_jess  <- elena_bothcodes %>%  mutate(check = ifelse(elena_bothcodes$frs_id %in% jess_bothcosedes$REGISTRY_ID, "both", "elena"))
subset_elena <- elena_not_jess %>% subset(check == "elena")
jess_not_elens  <- jess_bothcosedes %>%  mutate(check = ifelse(jess_bothcosedes$REGISTRY_ID %in% elena_bothcodes$frs_id, "both", "jess"))
subset_jess <- jess_not_elens %>%  subset(check == "jess")

```

```{r from Jess final data}

#use Jess final file
final_jess <- read.csv("FRS_FINAL.csv")
names(final_jess)
final_jess <- final_jess[, c(1:3, 6:11, 17,18)]
final_jess <- distinct(final_jess) #39962 

elena_bothcodes <- FRS_bothCodes
names(elena_bothcodes)
elena_bothcodes <- elena_bothcodes[,c (1:9)]
elena_bothcodes <- distinct(elena_bothcodes)

length(unique(jess_bothcosedes$REGISTRY_ID)) #39981
length(unique(elena_bothcodes$frs_id)) #43517

check_by_join <- full_join(elena_bothcodes, jess_bothcosedes, by = c("frs_id" = "REGISTRY_ID"))

elena_not_jess  <- elena_bothcodes %>%  mutate(check = ifelse(elena_bothcodes$frs_id %in% jess_bothcosedes$REGISTRY_ID, "both", "elena"))
subset_elena <- elena_not_jess %>% subset(check == "elena")
jess_not_elens  <- jess_bothcosedes %>%  mutate(check = ifelse(jess_bothcosedes$REGISTRY_ID %in% elena_bothcodes$frs_id, "both", "jess"))
subset_jess <- jess_not_elens %>%  subset(check == "jess")

```

```{r Creat a Master list}
list.files()


#use the state single file downloaded from https://www.epa.gov/frs/epa-frs-facilities-state-single-file-csv-download by clicking on CA. A single CSV will be downloaded
single_frs <- read.csv("STATE_SINGLE_CA.CSV")

#filter using determined frs_ids: Elena's file encompasses all FRS, one facility comes from Jess's list (is this a bug? ask Matt why the facility 110033614774 has NAICS in queried downloads, but does not have it in the upload whole files)

Master <- single_frs %>%  filter(REGISTRY_ID %in% elena_not_jess$frs_id | REGISTRY_ID == 110033614774)

names(Master)
Master <- clean_names(Master)
unique(Master$federal_facility_code)

##Master_tribalLend <- Master %>%  filter(tribal_land_code == "Yes")
#Master_mexicoBorder <- Master %>%  filter(us_mexico_border_ind == "Yes")

#Master <- Master %>% filter(federal_facility_code != "Yes") #this is a file with all except federal
#Master <- Master %>%  filter(tribal_land_code != "Yes")

#2-digit HUC first-level (region)
#4-digit HUC second-level (subregion)
#6-digit HUC third-level (accounting unit)
#8-digit HUC fourth-level (cataloguing unit)
#10-digit HUC fifth-level (watershed)
#12-digit HUC sixth-level (subwatershed) 

names(Master)

Master <- Master [,c(2:4, 6:7, 9, 12:16, 18,19,21,23,25:27,32:34)]
#Master_1 <- Master %>%  filter(naics_codes == "" & sic_codes == "") #CA-single file does not have NAICS code that appear on the file from ZIP, why?
vis_miss(Master, warn_large_data = F)
```


```{r}
Master <- Master %>%  mutate_all(na_if,"")
vis_miss(Master, warn_large_data = F)


Master$primary_name <- str_to_title(Master$primary_name)
Master$location_address <- str_to_title(Master$location_address)
Master$city_name <- str_to_title(Master$city_name)
Master$county_name <- str_to_title(Master$county_name)

Master_by_EG <- distinct(Master)
write.csv(Master_by_EG, "Master_by_EG.csv")

Master_by_JR <- read.csv("Master_by_JR.csv")
compare  <- Master_by_JR %>%  mutate(check = ifelse(Master_by_JR$RegistryID %in% Master_by_EG$registry_id, "both", "jess"))
table(compare$check)


```

